---
output:
  html_document: default
  pdf_document: default
---

```{r}
library(FNN)
library(caret)
library(ISLR)
library(dummies)
library(e1071)
library(tidyverse)
library(factoextra) 
library(class)
library(stats)
library(dplyr)
library(ggvis)
library(ggplot2)
library(pROC)
library(GGally)
library(flexclust)
library(imputeTS)
```

# Questions 1
```{r}
setwd("~/Desktop/Machine Learning 64060")
BathSoap <- read.csv("BathSoap.csv")
summary(BathSoap)
```


```{r}
Soap_Data <- BathSoap[20:46] %>% mutate_each(funs(as.numeric(gsub("%", "", ., fixed = TRUE))/100))
Soap <- cbind(BathSoap[1:19],Soap_Data)

Behavior<-Soap[,12:31]
```

```{r}
volume <- function(x){
return(x*Behavior$Total.Volume)
}
vol<-as.data.frame(lapply(Behavior[9:20],volume))
Purchase_Behavior <- Behavior[,1:8]
Purchase_Behav <- cbind(Purchase_Behavior,vol)
head(Purchase_Behav)
```

```{r}
Purchase_Behav$max <- apply(Purchase_Behav[,12:19], 1, max)
head(Purchase_Behav)
```

```{r}
Soap_scaled <- scale(Purchase_Behav[c(1:8,20,21)])
head(Soap_scaled)
```

```{r}
wss <- (nrow(Soap_scaled)-1)*sum(apply(Soap_scaled,2,var))
wss
```


```{r}
for (i in 2:15) 
  wss[i] <- sum(kmeans(Soap_scaled, 
                       centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters", 
     ylab="Within groups sum of squares",
     main="Assessing the Optimal Number of Clusters with the Elbow Method",
     pch=19, cex=2)
```

#k = 3.
```{r}
set.seed(123)
k3 <- kmeans(Soap_scaled, centers = 3, nstart = 25) 
```

```{r}
set.seed(123)
plot(Soap_scaled, col =(k3$cluster) , 
     main="K = 3", 
     pch=16, cex=2)
```


```{r}
k3$centers
k3$size
```

```{r}
#Cluster 1, size = 63, is higher loyal group, following main brands with a bigger individual purchases, with median value. 

#Cluster 2, size = 202, is a median loyal group, have many different brands with higher value.

#Cluster 3(size = 335) is lower loyal group, but may be of the least interest because the customers with least value.
```

```{r}
set.seed(123)
P_Basis<-Soap[,c(14,20:22,32:36,45)]
volume2 <- function(x){
return(x*P_Basis$Total.Volume)
}
Pur_Basis<-as.data.frame(lapply(P_Basis[2:10],volume2))
```

```{r}
Basis_scaled <- scale(Pur_Basis)
head(Basis_scaled)
```

```{r}
set.seed(123)
wss <- (nrow(Basis_scaled)-1)*sum(apply(Basis_scaled,2,var))
wss

for (i in 2:15) 
  wss[i] <- sum(kmeans(Basis_scaled, 
                       centers=i)$withinss)
```


```{r}
plot(1:15, wss, type="b", xlab="Number of Clusters", 
     ylab="Within groups sum of squares",
     main="Assessing the Optimal Number of Clusters with the Elbow Method",
     pch=20, cex=2)
```

```{r}
set.seed(123)
k4 <- kmeans(Basis_scaled, centers = 4, nstart = 25) 
```

```{r, fig.height=7, fig.width=12}
set.seed(123)
plot(Basis_scaled, col =(k4 $cluster) , 
     main="K-Means with 3 clusters", 
     pch=20, cex=2)
```


#Cluster 1 = 100, high loyal group, is notable for its responsiveness to price category 1, 2 and 4 and selling proposition 5 coupled with aversion to price categories 3 and selling proposition 14. 

#Cluster 2 = 47 better react to pricing category 3 promotions, and is responsive to selling proposition 14. 

#Cluster 3 = 453 has low react to ads, price categories 1, but it is not responsive to the 2 selling propositions.


```{r}
set.seed(123)
Both <-cbind(Soap_scaled, Basis_scaled)
k2_B <- kmeans(Both, centers = 2, nstart = 25) 
k2_B$centers
k2_B$size
```

```{r fig.height=7, fig.width=14}
cluster_Both <- c(1,2)
Both_clusters <- cbind(cluster_Both, k2_B$centers)


ggparcoord((Both_clusters),
           columns = 1:10, groupColumn = 1, 
           showPoints = TRUE, 
           alphaLines = 0.3 
)
```

```{r}
set.seed(123)
Demo <- Soap[2:11]
demo_scaled <- scale(Demo)
Both_Demo <- cbind(demo_scaled,Both)
```


```{r, fig.height=5, fig.width=12}
k2_Both_Demo <- kmeans(Both_Demo, centers = 2, nstart = 25) 
k2_Both_Demo$centers
k2_Both_Demo$size
```


```{r, fig.height=5, fig.width=12}
boxplot(Both_Demo)
```
#The two clusters are separated on almost all variables, Avg Price being an important exception.
Cluster1 = 69, is a more loyal cluster, with lower socioeconomic status and affluence.
```{r fig.height=5, fig.width=12}
# Question 2.
set.seed(123)
#Scree Plot - Check for the optimal number of clusters
wss2 <- (nrow(Both_Demo)-1)*sum(apply(Both_Demo,2,var))
wss2

for (i in 1:15) 
  wss2[i] <- sum(kmeans(Both_Demo, 
                       centers=i)$withinss)
```


```{r fig.height=5, fig.width=12}
plot(1:15, wss2, type="b", xlab="Number of Clusters", 
     ylab=" Groups Sum of Squares",
     main="Optimal Number of Clusters",
     pch=19, cex=2)
```

```{r fig.height=8, fig.width=21}
# k = 3, number of restarts = 25

set.seed(123)
k4_BothD <- kmeans(Both_Demo, centers = 4, nstart = 25) 
k4_BothD$centers
k4_BothD$size

cluster_Both_Demo4 <- c(1,2,3,4)
Both_Demo_clusters4 <- cbind(cluster_Both_Demo4, k4_BothD$centers)
```


```{r fig.height=8, fig.width=21}
ggparcoord((Both_Demo_clusters4),
           columns = 1:30, groupColumn = 1, 
           showPoints = TRUE, 
           alphaLines = 0.3 
)
```
#Cluster 1, n = 73 
#Itis characterized by low volume, loya, and pretty sensitive to ads and discount.
#(responsive to cat. 1, but not to  cat.2 and cat.3), have a smaller family size, and unmoved by selling proposition. 
#Demographically, it is affluent, of high socio-economic status.


#Cluster 1, n = 173 
#It is distinguished mostly by the purchase behavior variables. 
#It is lower brand loyalty group with high value, volume and frequency. The brand switching seems to be intrinsic. 
#This group is not particularly responsive to promotions, pricing or selling propositions.
#Demographically it is relatively affluent and educated.


#Cluster 3, n = 260
#I is a "gray" cluster, it is not characterized by very different values across all variables, but is responsive to price category 2 and selling proposition 5. 
#Demographically it is relatively affluent and well education.


#Cluster 4, n = 94 
#stands out in both groups of variables,it has high loyalty, low value and price per purchase
#Very different response to price (didn't responsive to categories 1, 2 and 4, but highly responsive to category 3), and selling proposition (didn't responsive to #5, highly responsive to #14).
#Demographically it has lower affluence and education level.

```{r}
#Question 3
set.seed(321)
k2_Model <- kcca(Both, k = 2, kccaFamily("kmeans")) 
k2_Model
pred <- predict(k2_Model, Both)
cluster_data <- data.frame(cluster = pred)
cluster_Demo <- cbind(cluster_data,Demo)

cluster_Demo$cluster <- ifelse(cluster_Demo$cluster==1,1,0)
head(cluster_Demo)
```


```{r}
cluster_Demo$cluster <- as.factor(cluster_Demo$cluster)
str(cluster_Demo)
```

```{r}
model <- glm(cluster~.,family="binomial", data=cluster_Demo)

summary(model)
```

```{r}
Probability <- predict(model, cluster_Demo, type="response")

Probability
```


```{r fig.height=6, fig.width=12}
Predictions <- ifelse(Probability > 0.35, 1, 0)
head(Probability)

table(Predictions, cluster_Demo$cluster)
mean(Predictions == cluster_Demo$cluster)

roc(cluster_Demo$cluster, Probability)
```


```{r fig.height=6, fig.width=12}
plot.roc(cluster_Demo$cluster,Probability)
```
